---
title: "assignment_7"
author: "Wyatt"
date: "2023-03-14"
output: github_document
---

```{r}
library(tidyverse)
library(knitr)
library(dslabs)
```

<br>

## Excercise: 2016 election result and polling

For this exercise, we will explore the result of the 2016 US
presidential election as well as the polling data. We will use the
following three datasets in the `dslabs` package, and use `join`
function to connect them together. As a reminder, you can use `?` to
learn more about these datasets.

- `results_us_election_2016`: Election results (popular vote) and
  electoral college votes from the 2016 presidential election.

- `polls_us_election_2016`: Poll results from the 2016 presidential
  elections.

- `murders`: Gun murder data from FBI reports. It also contains the
  population of each state.

We will also use [this
dataset](https://raw.githubusercontent.com/kshaffer/election2016/master/2016ElectionResultsByState.csv)
to get the exact numbers of votes for question 3.

<br>

### Question 1. What is the relationship between the population size and the number of electoral votes each state has?

**1a.** Use a `join` function to combine the `murders` dataset, which
contains information on population size, and the
`results_us_election_2016` dataset, which contains information on the
number of electoral votes. Name this new dataset `q_1a`, and show its
first 6 rows.

```{r}
q_1a <- murders %>% 
  left_join(results_us_election_2016)

q_1a %>% 
  head()

```
<br> <br>

**1b.** Add a new variable in the `q_1a` dataset to indicate which
candidate won in each state, and remove the columns `abb`, `region`, and
`total`. Name this new dataset `q_1b`, and show its first 6 rows.

```{r}
q_1b <- q_1a %>% 
  mutate(winner = ifelse(clinton>trump, "clinton", "trump")) %>% 
  subset(select = -c(abb, region, total))

q_1b %>% 
  head()

```

<br> <br>

**1c.** Using the `q_1b` dataset, plot the relationship between
population size and number of electoral votes. Use color to indicate who
won the state. Fit a straight line to the data, set its color to black,
size to 0.1, and turn off its confidence interval.

```{r}
q_1b %>% 
  ggplot(mapping = aes(x = population, y = electoral_votes)) +
  geom_point(mapping = aes(color = winner)) +
  geom_smooth(mapping = aes(x = population, y = electoral_votes), method = lm, se = FALSE, color = "black", size = 0.1)
```


### Question 2. Would the election result be any different if the number of electoral votes is exactly proportional to a state’s population size?

**2a.** First, convert the `q_1b` dataset to longer format such that the
`population` and `electoral_votes` columns are turned into rows as shown
below. Name this new dataset `q_2a`, and show its first 6 rows.

```{r}
q_2a <- q_1b %>% 
  pivot_longer(c( population, electoral_votes), names_to = 'metric', values_to = 'value')

q_2a %>% 
  head()
```
<br> <br>

**2b.** Then, sum up the number of electoral votes and population size
across all states for each candidate. Name this new dataset `q_2b`, and
print it as shown below.

```{r}
q_2b <- q_2a %>% 
  group_by(metric, winner) %>% 
  summarize(value = sum(value))

q_2b

```
<br> <br>

**2c.** Use the `q_2b` dataset to contruct a bar plot to show the final
electoral vote share under the scenarios of **1)** each state has the
number of electoral votes that it currently has, and **2)** each state
has the number of electoral votes that is exactly proportional to its
population size. Here, assume that for each state, the winner will take
all its electoral votes.

```{r}
q_2b %>% 
  ggplot(mapping = aes(x = metric, y = value, fill = winner)) +
  geom_col(position = "fill")
```

<br> <br>

### Question 3. What if the election was determined by popular votes?

**3a.** First, from [this dataset on
GitHub](https://raw.githubusercontent.com/kshaffer/election2016/master/2016ElectionResultsByState.csv),
calculate the number of popular votes each candidate received as shown
below. Name this new dataset `q_3a`, and print it. <br>

*Note: all candidates other than Clinton and Trump are included in
`others` as shown below.*

*Hint: `pivot_longer()` may be useful in here.*

```{r}
q_3a <- dataset3 %>% 
  summarise(clinton = sum(clintonVotes),trump = sum(trumpVotes), others = sum(totalVotes - clintonVotes - trumpVotes)) %>% 
  pivot_longer(cols = c("clinton", "trump", "others"), names_to = "winner", values_to = "value") %>% 
  mutate(metric = "popular_votes") %>% 
  select(metric, winner, value)

q_3a
```
<br> <br>

**3b.** Combine the `q_2b` dataset with the `q_3a` dataset. Call this
new dataset `q_3b`, and print it as shown below.

```{r}
q_3b <- q_2b %>% 
  full_join(q_3a)

q_3b
```
<br> <br>

**3c.** Lastly, use the `q_3b` dataset to contruct a bar plot to show
the final vote share under the scenarios of **1)** each state has the
number of electoral votes that it currently has, **2)** each state has
the number of electoral votes that is exactly proportional to its
population size, and **3)** the election result is determined by the
popular vote.

```{r}
```






